Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-02-15 scagnostics.Rmd
filling in some of the investigation

@@

## Scagnostic discreteness

@@

@@

# Modification to scagnostics

>>>>

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE, 
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
)
```

@@

The nine scagnostics measures are designed to be computed together, producing a matrix of values with dimensions corresponding to the number of all pairs of variables, and the number of scagnostics. If there are 100 variables, the dimension will be $4950x9$. It is typically used to extract the pairs of variables that have the highest scagnostic values, and thus considered to be the most interesting. It can also be used to do more in-depth exploratory analysis, and is interesting to study the distribution of the scagnostics resulting from the data, as a data set in itself. In a sense this the data's "fingerprint".

To use scagnostics for projection pursuit requires that each of the measures for functions that are relatively smooth over the space of all 2D projections, and that they can be computed rapidly enough that the display can be updated many times a second during optimisation.

## Smoothness

@@

XXX Probably use a simple benchmark data to illustrate the issues, maybe mtcars that comes with the scagnostics package

```{r load}
library(tourr)
library(tidyverse)
library(reshape2)
library(binostics) #https://github.com/uschiLaa/binostics
library(gridExtra)
library(wesanderson) #color palette
library(tictoc) #timer
library(mbgraphic) #Katrins package
library(dfoptim)
```

```{r util}
scagIndex <- function(scagType){
  function(mat){
    sR <- scagnostics.default(mat[,1],mat[,2])$s
    return(sR[scagType])
  }
}
```

```{r tourpath}
data(mtcars)
t1 <- save_history(mtcars, grand_tour(2), max=4)
t1full <- as.list(interpolate(t1))
n <- length(t1full)
sc <- NULL
for (i in 1:n) {
  dprj <- as.matrix(mtcars) %*% t1full[[i]]
  sc <- rbind(sc, scagnostics(dprj))
}
sc <- as.data.frame(sc)
sc$t <- 1:n
sc <- sc %>% gather(index, value, -t)
```

```{r plotscag, fig.width=6, fig.height=7, fig.cap="Short tour paths with scagnostics computed on projections. Most of the indices exhibit sharp jumps in scagnostic value."}
ggplot(sc, aes(x=t, y=value)) + geom_line() + 
  facet_wrap(~index, ncol=1, scales = "free_y")
```

## Effect of binning

@@

# Possible modifications
